---
title: "R code for some plots and results"
output: html_notebook
---

# Code to obtain the plots and results

Please, ensure having run the code in the Functions Section below before running the following blocks.

```{r}
corr_analysis(Data_Pre)
corr_analysis(Data_Post)
```

```{r}
IPR_fun_KS(Data_Pre)
IPR_fun_KS(Data_Post)

IPR_fun_JB(Data_Pre)
IPR_fun_JB(Data_Post)
```

```{r}
mp_plotting(Data_Pre)
mp_plotting(Data_Post)
```

```{r}
MP_and_null(Data_Pre)
MP_and_null(Data_Post)
```

```{r}
temporal_corr_indexes(Data_Pre, 100, 20)
temporal_corr_indexes(Data_Pre, 100, 40)
temporal_corr_indexes(Data_Pre, 100, 60)
temporal_corr_indexes(Data_Pre, 100, 80)
temporal_corr_indexes(Data_Pre, 100, 100)

temporal_corr_indexes(Data_Post, 100, 20)
temporal_corr_indexes(Data_Post, 100, 40)
temporal_corr_indexes(Data_Post, 100, 60)
temporal_corr_indexes(Data_Post, 100, 80)
temporal_corr_indexes(Data_Post, 100, 100)
```

```{r}
plot_grid(Data_Pre)
plot_grid(Data_Post)
```

```{r}
residuals_matrix <- linear_model_residuals(Data_Pre)

plot(density(cor(residuals_matrix)), main = paste("Distribution of correlation matrices of Data_Post")
     , col = "orange", xlab = "Cross-correlaion coefficient", type = "l")
lines(density(cor(Data_Pre)), col = "green")
legend("topright", legend = c("First eigenvector removed", "Correlation matrix"),
       col = c("orange", "green"), lwd = 2, lty = c(1, 1))

residuals_matrix <- linear_model_residuals(Data_Post)

plot(density(cor(residuals_matrix)), main = paste("Distribution of correlation matrices of Data_Post")
     , col = "orange", xlab = "Cross-correlaion coefficient", type = "l")
lines(density(cor(Data_Pre)), col = "green")
legend("topright", legend = c("First eigenvector removed", "Correlation matrix"),
       col = c("orange", "green"), lwd = 2, lty = c(1, 1))
```

```{r}
residuals_matrix <- linear_model_residuals(Data_Pre)
get_top_entries(residuals_matrix)
smallest_eigenvectors(residuals_matrix)
```

```{r}
residuals_matrix <- linear_model_residuals(Data_Post)
get_top_entries(residuals_matrix)
smallest_eigenvectors(residuals_matrix)
```

# Functions section

Necessary libraries

```{r}
# Load necessary libraries
library(readr)
library(stats)
library(ggplot2)
library(MASS)
```

The following code chunk loads the data.

```{r}

# Load Data and transpose
Data_Pre <- read_csv("Datasets/Data_PreCovid_20170101_20200109.csv", col_names = FALSE)
Data_Post <- read_csv("Datasets/Data_PostCovid_20200110_20221231.csv", col_names = FALSE)

Data_Pre <- t(Data_Pre)
Data_Post <- t(Data_Post)
```

The following code chunk is a function used by many of the subsequent functions.

```{r}
plot_heatmap <- function(corr_matrix) {
  
  # Convert the correlation matrix to a long-format data frame using melt
  corr_melted <- melt(corr_matrix)
  
  # Number of rows in a square correlation matrix (assuming it is 45x45)
  n <- sqrt(nrow(corr_melted))  # Assuming the matrix is square and 45x45
  
  # Generate indices for adjusting Var1 and Var2
  j <- seq(from = 1, to = length(corr_melted$Var1), by = n)
  
  # Adjust Var1 and Var2 with appropriate indexing
  for (i in 1:length(j)) {
    if ((j[i] + (n - 1)) <= length(corr_melted$Var1)) {
      corr_melted$Var1[j[i]:(j[i] + (n - 1))] <- seq(1, n)
      corr_melted$Var2[j[i]:(j[i] + (n - 1))] <- rep(i, times = n)
    }
  }
  
  # Create the heatmap using ggplot2 with reversed y-axis
  ggplot(data = corr_melted, aes(x = Var1, y = Var2, fill = value)) +
    geom_tile() +
    scale_fill_gradient2(low = "blue", high = "red", mid = "white", 
                         midpoint = 0, limit = c(-1, 1), space = "Lab", 
                         name = "Correlation", na.value = "green") +  # Ensure NA values are white
    scale_y_reverse() +  # Reverse the y-axis
    theme_minimal() +
    labs(title = "Heatmap of Correlation Matrix",
         x = "Variables",
         y = "Variables") +
    theme(axis.text.x = element_text(angle = 45, vjust = 1, hjust = 1, size = 14),  # Increased font size
          axis.text.y = element_text(size = 14),  # Increased font size
          axis.title.x = element_text(size = 16),  # Font size for x-axis label
          axis.title.y = element_text(size = 16))  # Font size for y-axis label
}
```

Function to obtain the heatmap of the correlation matrix of a dataset.

```{r}
corr_analysis <- function(Dataset){
  corr_matrix <- cor(Dataset)
  print(plot_heatmap(corr_matrix = corr_matrix))
}
```

Two functions to check gaussianity via the Kolmogorov-Smirnov test and the Jarque-Baret hypothesis test. Load it as it is used by other functions.

```{r}
# Function to check for gaussianity in the eigenvectors
# Function to apply the Kolmogorov-Smirnov test and return non-Gaussian eigenvectors and their positions
check_gaussianity <- function(Dataset, alpha = 0.05) {
  non_gaussian <- list()  # To store non-Gaussian eigenvectors
  indices <- vector()     # To store corresponding indices
  
  corr_matrix <- scaled(Dataset = Dataset)
  
  eigenvectors <- eigen(corr_matrix, symmetric = T)$vectors
  
  for (i in 1:ncol(eigenvectors)) {
    eigenvector <- eigenvectors[, i]
    
    # Perform the KS test comparing the eigenvector to a normal distribution
    ks_test <- ks.test(eigenvector, "pnorm", mean = mean(eigenvector), sd = sd(eigenvector))
    
    # If the p-value is less than alpha, the eigenvector is considered non-Gaussian
    if (ks_test$p.value < alpha) {
      non_gaussian[[length(non_gaussian) + 1]] <- eigenvector
      indices <- c(indices, i)  # Save index
    }
  }
  return(list(non_gaussian_vectors = non_gaussian, positions = indices))
}

JB_gaussianity <- function(Dataset, alpha = 0.01) {
  non_gaussian <- list()  # To store non-Gaussian eigenvectors
  indices <- vector()     # To store corresponding indices
  
  # Compute the correlation matrix
  corr_matrix <- cor(Dataset)
  
  # Perform eigen decomposition
  eigenvectors <- eigen(corr_matrix, symmetric = TRUE)$vectors
  
  # Iterate over each eigenvector
  for (i in 1:ncol(eigenvectors)) {
    eigenvector <- eigenvectors[, i]
    
    # Perform the Jarque-Bera test
    jb_test <- jarque.bera.test(eigenvector)
    
    # If the p-value is less than alpha, the eigenvector is considered non-Gaussian
    if (jb_test$p.value < alpha) {
      non_gaussian[[length(non_gaussian) + 1]] <- eigenvector
      indices <- c(indices, i)  # Save index
    }
  }
  
  # Return the list of non-Gaussian eigenvectors and their indices
  return(list(non_gaussian_vectors = non_gaussian, positions = indices))
}
```

Two functions to obtain two null models: a GOE and a shuffled data. Load it as it is used by other functions.

```{r}
generate_GOE <- function(m, n) {
  # Create a matrix with normally distributed random numbers
  GOE_matrix <- matrix(rnorm(n * m), nrow = m, ncol = n)
  
  return(GOE_matrix)
}

randomize_null <- function(Dataset) {
  
  n_rows <- nrow(Dataset)
  n_cols <- ncol(Dataset)
  
  # Flatten the Dataset into a vector
  Dataset_vector <- as.vector(Dataset)
  
  # Randomize the positions
  order <- sample(1:length(Dataset_vector), size = length(Dataset_vector), replace = FALSE)
  
  # Create a new randomized matrix with the same dimensions
  Dataset_randomized <- matrix(Dataset_vector[order], nrow = n_rows, ncol = n_cols)
  
  return(Dataset_randomized)
}
```

Function to calculate the IPR along with the Kolmogorov-Smirnov test

```{r}
IPR_fun_KS <- function(Dataset) {
  
  # Number of variables (N) and observations (M) for MP-Law parameters
  N <- ncol(Dataset)  # Number of variables
  M <- nrow(Dataset)  # Number of observations
  Q <- N / M          # Shape parameter (N / M)
  
  # Calculate the range for the Marchenko-Pastur distribution
  lambda_min <- (1 - sqrt(Q))^2
  lambda_max <- (1 + sqrt(Q))^2
  
  # Fit the Marchenko-Pastur distribution to get the fitted bounds
  corr_matrix <- cor(Dataset)
  eigenvalues <- eigen(corr_matrix, symmetric = TRUE)$values
  

  if (Dataset == Data_Pre){
  lambda_plus_fitted <- 1.029  
  lambda_minus_fitted <- 0.1138 
  } else{
  lambda_plus_fitted <- 0.872  
  lambda_minus_fitted <- 0.083 
  }
  
  vectors <- eigen(corr_matrix)$vectors
  values <- eigenvalues
  
  IPRs <- vector(mode = "numeric", length = 98)
  
  for(i in 1:nrow(corr_matrix)){
    IPRs[i] <- sum(vectors[,i]^4)
  }
  
  non_gaussian <- check_gaussianity(Dataset)
  
  # Highlight and plot non-Gaussian eigenvalues
  non_gaussian_positions <- non_gaussian[["positions"]]
  
  data_plot <- data.frame(eigenvalues = values, IPR = IPRs, gaussian_positions = rep("blue", times = 98))
  data_plot$gaussian_positions[non_gaussian_positions] <- "green" 
  
  # Plot the IPR against eigenvalues
  plot(x = log10(data_plot$eigenvalues), y = log10(data_plot$IPR), type = "b",
       xlab = "log(Eigenvalues)", ylab = "log(Inverse Participation Ratio)", col = data_plot$gaussian_positions,
       main = paste("Plot of IPR against eigenvalues of ", deparse(substitute(Dataset)), "using Kolmogorov-Smirnov test"), 
       xlim = c(-2, 1.5), ylim = c(-2, -0.2), cex.axis = 1.5, cex.lab = 1.5)
  
  # Highlight the Marchenko-Pastur distribution range (Gray Rectangle)
  rect(log10(lambda_min), -2, log10(lambda_max), 0, col = rgb(0.5, 0.5, 0.5, 1/4))
  abline(v = log10(lambda_min), col = "red")
  abline(v = log10(lambda_max), col = "red")
  
  # Add vertical pink lines for the fitted values (Fitted MP-Law)
  rect(log10(lambda_minus_fitted), -2, log10(lambda_plus_fitted), 0, col = alpha("pink", 0.5))
  abline(v = log10(lambda_plus_fitted), col = "pink", lty = 2, lwd = 2)  # Upper fitted bound
  abline(v = log10(lambda_minus_fitted), col = "pink", lty = 2, lwd = 2)  # Lower fitted bound
  
  # Draw the black lines connecting the points
  lines(log10(data_plot$eigenvalues), log10(data_plot$IPR), col = "black", lty = 1)
  
  # Add a legend with information about the bounds
  legend("topright", 
         legend = c("Associated to Non-Gaussian eigenvectors", 
                    "Associated to Gaussian eigenvectors", 
                    "MP-Law bounds", 
                    "Fitted MP-Law bounds"), 
         col = c("green", "blue", rgb(0.5, 0.5, 0.5, 1/4), alpha("pink", 0.5)), 
         pch = c(16, 16, 15, 15),  # Use filled squares for rectangles
         pt.cex = 2,
         lty = c(NA, NA, NA, 2),  # Line type for pink bounds
         lwd = c(NA, NA, NA, 2))  # Line width for pink bounds
}
```

Function to calculate the IPR along with the Jarque-Baret hypothesis test

```{r}
IPR_fun_JB <- function(Dataset){
  
  # Number of variables (N) and observations (M) for MP-Law parameters
  N <- ncol(Dataset)  # Number of variables
  M <- nrow(Dataset)  # Number of observations
  Q <- N / M          # Shape parameter (N / M)
  
  # Calculate the range for the Marchenko-Pastur distribution
  lambda_min <- (1 - sqrt(Q))^2
  lambda_max <- (1 + sqrt(Q))^2
  
  if (Dataset == Data_Pre){
  lambda_plus_fitted <- 1.029  
  lambda_minus_fitted <- 0.1138 
  } else{
  lambda_plus_fitted <- 0.872  
  lambda_minus_fitted <- 0.083 
  }
  
  # Compute the correlation matrix
  corr_matrix <- cor(Dataset)
  
  # Perform eigen decomposition
  eigen_decomp <- eigen(corr_matrix)
  vectors <- eigen_decomp$vectors
  values <- eigen_decomp$values
  
  # Calculate IPR for each eigenvector
  IPRs <- vector(mode = "numeric", length = N)
  for(i in 1:N){
    IPRs[i] <- sum(vectors[, i]^4)
  }
  
  # Check Gaussianity using the Jarque-Bera test (updated function)
  non_gaussian <- JB_gaussianity(Dataset = Dataset)
  
  # Extract non-Gaussian eigenvector positions
  non_gaussian_positions <- non_gaussian[["positions"]]
  
  # Create a data frame for plotting
  data_plot <- data.frame(eigenvalues = values, IPR = IPRs, gaussian_positions = rep("blue", times = N))
  
  # Mark non-Gaussian eigenvectors in green
  data_plot$gaussian_positions[non_gaussian_positions] <- "green" 
  print(data_plot)
  # Plot IPR against eigenvalues with log scale
  plot(x = log10(data_plot$eigenvalues), y = log10(data_plot$IPR), type = "b",
       xlab = "log(Eigenvalues)", ylab = "log(Inverse Participation Ratio)", col = data_plot$gaussian_positions,
       main = paste("Plot of IPR against eigenvalues of", deparse(substitute(Dataset)), "using Jarque-Bera test"), 
       xlim = c(-2, 1.5), ylim = c(-2, -0.2), cex.lab = 1.5, cex.axis = 1.5)
  
  # Add vertical pink lines for the fitted values
  rect(log10(lambda_plus_fitted), -2, log10(lambda_minus_fitted), 0, col = alpha("pink",0.5), )
  abline(v = log10(lambda_plus_fitted), col = "pink", lty = 2, lwd = 2)  # Upper fitted bound
  abline(v = log10(lambda_minus_fitted), col = "pink", lty = 2, lwd = 2)  # Lower fitted bound
  
  # Draw the black lines connecting the points
  lines(log10(data_plot$eigenvalues), log10(data_plot$IPR), col = "black", lty = 1)
  
  # Highlight the Marchenko-Pastur distribution range
  rect(log10(lambda_min), -2, log10(lambda_max), 0, col = rgb(0.5, 0.5, 0.5, 1/4))
  abline(v = log10(lambda_min), col = "red")
  abline(v = log10(lambda_max), col = "red")
  
  # Add a legend with information about the bounds
  legend("topright", 
         legend = c("Associated to Non-Gaussian eigenvectors", 
                    "Associated to Gaussian eigenvectors", 
                    "MP-Law bounds", 
                    "Fitted MP-Law bounds"), 
         col = c("green", "blue", rgb(0.5, 0.5, 0.5, 1/4), alpha("pink", 0.5)), 
         pch = c(16, 16, 15, 15),  # Use filled squares for rectangles
         pt.cex = 2,
         lty = c(NA, NA, NA, 2),  # Line type for pink bounds
         lwd = c(NA, NA, NA, 2))  # Line width for pink bounds
}
```

Function to calculate the MP-Law on a dataset

```{r}
mp_density <- function(lambda, Q) {
  # Calculate the range for the Marchenko-Pastur distribution
  lambda_min <- (1 - sqrt(Q))^2
  lambda_max <- (1 + sqrt(Q))^2
  sqrt_part <- sqrt((lambda_max - lambda) * (lambda - lambda_min))
  return((1 / (2 * pi * Q * lambda)) * (sqrt_part))
}
```

This function produces the histogram of the eigenvalues with the MP-law and GOE histogram of eigenvalues overplotted

```{r}
mp_plotting <- function(Dataset){
  
  # Scale and compute the correlation matrix
  corr_matrix <- cor(Dataset)
  
  # Get eigenvalues from the correlation matrix
  eigenval <- eigen(corr_matrix, symmetric = TRUE)$values
  
  # Plot histogram of eigenvalues from the dataset
  hist(eigenval, breaks = 200, probability = TRUE, 
       main = paste("Histogram of Eigenvalues with MP Law", deparse(substitute(Dataset))),
       xlab = "Eigenvalues", col = "lightblue", ylim = c(0, 1.5), xlim = c(0,10), cex.axis = 1.5)
  
  # Number of variables (N) and observations (M) for MP-Law parameters
  N <- ncol(Dataset)  # Number of variables
  M <- nrow(Dataset)  # Number of observations
  Q <- N / M          # Shape parameter (N / M)
  
  # Calculate the range for the Marchenko-Pastur distribution
  lambda_min <- (1 - sqrt(Q))^2
  lambda_max <- (1 + sqrt(Q))^2
  
  # Sequence for lambda and MP distribution values
  lambda_seq <- seq(lambda_min, lambda_max, length.out = 1000)
  mp_values <- sapply(lambda_seq, mp_density, Q = Q)
  
  # Add MP Law curve to the plot
  lines(lambda_seq, mp_values, col = "red", lwd = 2, lty = 2)
  
  # Generate a GOE matrix of size N x N (same size as correlation matrix)
  #GOE_matrix <- generate_GOE(m = M, n = N) 
  
  #GOE_corr <- cor(GOE_matrix)
  
  #GOE_eigen <- eigen(GOE_corr)$values
  
  # Flatten the GOE matrix to a vector to get all entries
  #GOE_entries <- as.vector(GOE_eigen)
  
  # Add density curve for the GOE matrix
  #lines(density(GOE_entries), col = "green", lwd = 2)
  
  # Add legend
  legend("topright", legend = c("Eigenvalues", "MP-Law"),
         col = c("lightblue", "red"), lwd = 2, lty = c(1, 2, 1))
}
```

Function to compute the MP against the shuffled data

```{r}
MP_and_null <- function(Dataset){
  # Number of variables (N) and observations (M) for MP-Law parameters
  N <- ncol(Dataset)  # Number of variables
  M <- nrow(Dataset)  # Number of observations
  Q <- N / M          # Shape parameter (N / M)
  
  GOE_matrix <- generate_GOE(m = M, n = N) 
  
  GOE_corr <- cor(GOE_matrix)
  
  GOE_eigen <- eigen(GOE_corr)$values
  
  # Flatten the GOE matrix to a vector to get all entries
  GOE_entries <- as.vector(GOE_eigen)
  
  control <- randomize_null(Dataset)
  
  control <- eigen(cor(control))$values
  
  
  # Calculate the range for the Marchenko-Pastur distribution
  lambda_min <- (1 - sqrt(Q))^2
  lambda_max <- (1 + sqrt(Q))^2
  
  # Sequence for lambda and MP distribution values
  lambda_seq <- seq(lambda_min, lambda_max, length.out = 1000)
  mp_values <- sapply(lambda_seq, mp_density, Q = Q)
  
  plot(y = density(GOE_entries)$y, x = density(GOE_entries)$x, type = "l", ylim = c(0, 1.2), main = paste("Comparison of histograms of",deparse(substitute(Dataset)))
       , ylab = "density", xlab = "Eigenvalues", cex.lab = 1.7, cex.axis = 1.7)
  
  # Add density curve for the GOE matrix
  lines(lambda_seq, mp_values, col = "red", lwd = 2, lty = 2)
  lines(density(control), col = "orange", lwd = 2)
  
  #Add legend
  legend("topright", legend = c("Control shuffle", "MP-Law", "GOE Entries"),
         col = c("orange", "red", "black"), lwd = 2, lty = c(1, 2, 1))
  
}
```

Function to track change in the top 10 entries.

```{r}
# Tracking changes in top 10 entries
temporal_corr_indexes <- function(Dataset, window_size, step_width) {
  
  # Number of variables (N) and observations (M) for MP-Law parameters
  N <- ncol(Dataset)  # Number of variables
  M <- nrow(Dataset)  # Number of observations
  Q <- N / M          # Shape parameter (N / M)
  
  # Maximum and minimum eigenvalues based on MP-Law
  lambda_max <- (1 + sqrt(Q))^2
  
  # Eigenvalue decomposition of the correlation matrix of the dataset
  eigenvalues <- eigen(cor(Dataset))$values
  
  # Find positions of eigenvalues outside the [lambda_min, lambda_max] range
  positions <- which(eigenvalues > lambda_max)
  num_positions <- length(positions)
  
  # Calculate the number of cycles
  cycles <- floor((nrow(Dataset) - (window_size + 1)) / step_width)
  
  # Initialize a vector to track changes for each eigenvector
  change_counts <- matrix(NA, nrow = cycles, ncol = num_positions)
  
  for (i in 1:cycles) {
    # Adjust the sliding window calculation
    start_A <- i
    end_A <- start_A + window_size
    start_B <- start_A + step_width
    end_B <- start_B + window_size
    
    if (end_B > nrow(Dataset)) break
    
    Data_A <- Dataset[start_A:end_A, ]
    Data_B <- Dataset[start_B:end_B, ]
    
    # Compute correlation matrices
    Data_A <- cor(Data_A)
    Data_B <- cor(Data_B)
    
    # Eigen decomposition
    eigen_A <- eigen(Data_A, symmetric = TRUE)
    eigen_B <- eigen(Data_B, symmetric = TRUE)
    
    # Select only the eigenvectors based on the 'positions'
    vectors_A <- eigen_A$vectors[, positions, drop = FALSE]  # Ensure matrix format
    vectors_B <- eigen_B$vectors[, positions, drop = FALSE]  # Ensure matrix format
    
    # Track changes in the top 10 entries for each eigenvector
    for(n in 1:num_positions){
      # Get the positions of the top 10 entries in absolute value
      top_10_A <- order(abs(vectors_A[, n]), decreasing = TRUE)[1:10]
      top_10_B <- order(abs(vectors_B[, n]), decreasing = TRUE)[1:10]
      # Count how many positions are different between vectors_A and vectors_B
      change_counts[i,n] <- sum(!top_10_A %in% top_10_B)
    }
  }
  change_counts <- change_counts/10
  # Convert matrix to data frame for labeling purposes
  change_counts_df <- as.data.frame(change_counts)
  
  # Create a boxplot for each column
  boxplot(change_counts_df,
          xlab = "Eigenvectors",
          ylab = "change",
          col = "lightblue",
          border = "darkblue",
          notch = F,
          ylim = c(0,1),
          cex.axis = 1.5,
          cex.lab = 1.5)
}
```

This function provides the heatmap of the eigenvectors to the right of the bulk

```{r}
#Heatmaps with cropping
temporal_corr_cropped <- function(Dataset, window_size, step_width) {
  
  # Number of variables (N) and observations (M) for MP-Law parameters
  N <- ncol(Dataset)  # Number of variables
  M <- nrow(Dataset)  # Number of observations
  Q <- N / M          # Shape parameter (N / M)
  
  # Maximum and minimum eigenvalues based on MP-Law
  lambda_max <- (1 + sqrt(Q))^2
  #lambda_min <- (1 - sqrt(Q))^2
  
  # Eigenvalue decomposition of the correlation matrix of the dataset
  eigenvalues <- eigen(cor(Dataset))$values
  
  # Find positions of eigenvalues outside the [lambda_min, lambda_max] range
  positions <- which(eigenvalues > lambda_max)
  num_positions <- length(positions)
  
  # Initialize the temporal average matrix with the correct dimensions
  temporal_average <- matrix(0, nrow = num_positions, ncol = num_positions)
  
  # Set row and column names based on positions
  rownames(temporal_average) <- positions
  colnames(temporal_average) <- positions
  
  # Calculate the number of cycles
  cycles <- floor((nrow(Dataset) - (window_size + 1)) / step_width)
  
  for (i in 1:cycles) {
    # Adjust the sliding window calculation
    start_A <- i
    end_A <- start_A + window_size
    start_B <- start_A + step_width
    end_B <- start_B + window_size
    
    if (end_B > nrow(Dataset)) break
    
    Data_A <- Dataset[start_A:end_A, ]
    Data_B <- Dataset[start_B:end_B, ]
    
    # Compute correlation matrices
    Data_A <- cor(Data_A)
    Data_B <- cor(Data_B)
    
    # Eigen decomposition
    eigen_A <- eigen(Data_A, symmetric = TRUE)
    eigen_B <- eigen(Data_B, symmetric = TRUE)
    
    # Select only the eigenvectors based on the 'positions'
    vectors_A <- eigen_A$vectors[, positions, drop = FALSE]  # Ensure matrix format
    vectors_B <- eigen_B$vectors[, positions, drop = FALSE]  # Ensure matrix format
    
    # Calculate temporal correlation matrix
    temporal <- t(vectors_A) %*% vectors_B
    
    # Accumulate the temporal correlation values
    temporal_average <- temporal_average + temporal
  }
  
  # Average the temporal correlations across all cycles
  temporal_average <- temporal_average / i
  
  # Final heatmap with correct labels
  plot_heatmap(temporal_average)
}

# Assuming 'temporal_corr_cropped' returns a ggplot object
plot_grid <- function(Dataset) {
  
  # Initialize an empty list to store the plots
  plot_list <- list()
  
  # Loop through different step_width values (20, 40, 60, 80, 100)
  for (i in 1:5) {
    step_width <- i * 20
    # Call the function and store the result in the list
    plot_list[[i]] <- temporal_corr_cropped(Dataset, window_size = 100, step_width = step_width)
  }
  
  # Arrange the 5 plots in a grid
  grid.arrange(grobs = plot_list, ncol = 2)
}
```

Function to compute the linear model

```{r}
linear_model_residuals <- function(Dataset) {
  
  N <- nrow(Dataset)
  M <- ncol(Dataset)
  scaled_data <- scale(Dataset)
  corr_matrix <- cor(Dataset)
  
  first_eigenvector <- eigen(corr_matrix)$vectors[, 1]
  
  G_t <- vector(mode = "numeric", length = N)
  M_t <- vector(mode = "numeric", length = N)
  
  # Initialize a matrix to store residuals 
  residuals_matrix <- matrix(nrow = N, ncol = M)
  
  for (j in 1:M) {
    for (i in 1:N) {
      M_t[i] <- t(first_eigenvector) %*% scaled_data[i, ]
      G_t[i] <- scaled_data[i, j]
    }
    
    # Fit the linear model and store residuals in the corresponding column
    residuals_matrix[, j] <- lm(G_t ~ M_t)$residuals
  }
  
  return(residuals_matrix)
}
```

```{r}
residuals_matrix <- linear_model_residuals(Data_Post)

plot(density(cor(residuals_matrix)), main = paste("Distribution of correlation matrices of Data_Post")
     , col = "orange", xlab = "Cross-correlaion coefficient", type = "l")
lines(density(cor(Data_Pre)), col = "green")
legend("topright", legend = c("First eigenvector removed", "Correlation matrix"),
       col = c("orange", "green"), lwd = 2, lty = c(1, 1))
```

Function to get the first 10 top entries from the eigenvectors to the right of the bulk

```{r}
get_top_entries <- function(residuals_matrix) {
  # Compute the correlation matrix of the residuals
  corr_residuals <- cor(residuals_matrix)
  
  # Perform eigen decomposition and extract the first 10 eigenvectors
  eigenvectors <- eigen(corr_residuals)$vectors[, 1:10]
  
  # Initialize a list to store the top 10 entries for each eigenvector
  top_entries <- list()
  
  # Loop over the first 10 eigenvectors
  for (i in 1:10) {
    # Get the absolute values of the eigenvector entries
    abs_entries <- abs(eigenvectors[, i])
    
    # Find the indices of the top 10 largest absolute values
    top_indices <- order(abs_entries, decreasing = TRUE)[1:10]
    
    # Store the indices and corresponding values in the list
    top_entries[[i]] <- list(indices = top_indices, values = eigenvectors[top_indices, i])
  }
  
  return(top_entries)
}
```

Function to get the first 2 top entries from the first 5 eigenvectors to the left of the bulk

```{r}
smallest_eigenvectors <- function(residuals_matrix){
  
  # Compute the correlation matrix of the residuals
  corr_residuals <- cor(residuals_matrix)
  
  # Perform eigen decomposition and extract the first 10 eigenvectors
  eigenvectors <- eigen(corr_residuals)$vectors
  eigenvectors <- eigenvectors[, tail(1:ncol(eigenvectors), 5)]
  
  # Initialize a list to store the top 10 entries for each eigenvector
  top_entries <- list()
  
  # Loop over the first 10 eigenvectors
  for (i in 1:ncol(eigenvectors)) {
    # Get the absolute values of the eigenvector entries
    abs_entries <- abs(eigenvectors[, i])
    
    # Find the indices of the top 2 largest absolute values
    top_indices <- order(abs_entries, decreasing = TRUE)[1:2]
    
    # Store the indices and corresponding values in the list
    top_entries[[i]] <- list(indices = top_indices, values = eigenvectors[top_indices, i])
  }
  
  return(top_entries)
}
```
