---
title: "R Notebook"
output: html_notebook
---


The following code chunk provides the necessary libraries
```{r}
# Load necessary libraries
library(readr)
library(stats)
library(ggplot2)
library(MASS)
```

The following code chunk loads the data.
```{r}

# Load Data and transpose
Data_Pre <- read_csv("Downloads/Data_PreCovid_20170101_20200109.csv", col_names = FALSE)
Data_Post <- read_csv("Downloads/Data_PostCovid_20200110_20221231.csv", col_names = FALSE)

Data_Pre <- t(Data_Pre)
Data_Post <- t(Data_Post)
```

The following code chunk is a function used by many of the subsequent functions.
```{r}
#A function to center and scale the data and compute the correlation matrix
scaled <- function(Dataset){
  centred <- scale(Dataset)
  correlation <- cov(centred)
}
```

The following code chunk is a function to obtain three plots out of the correlation matrix.
```{r}
#A function to obtain the corraltion matrix heatmap, histogram, and entries' boxplot 
corr_analysis <- function(Dataset){
  
  corr_matrix <- scaled(Dataset)
  heatmap(corr_matrix, Rowv = NA, Colv = NA)
  hist(corr_matrix, xlim = c(-1,1), breaks = 20, freq = T)
  abline(v=mean(corr_matrix), col = "red")
  abline(v=median(corr_matrix))
  boxplot(corr_matrix)
  abline(h = mean(corr_matrix), col="red")
}
```

```{r}
corr_analysis(Dataset = Data_Pre)
```

```{r}
corr_analysis(Dataset = Data_Post)
```


Check_gaussianity is a function used by the function IPR_fun. IPR_fun function plots the log of the Inverse partecipation Ratio against the log the eigenvalues. The colors show whether the corresponding eigenvector is Gaussian or not. 
```{r}
# Function to check for gaussianity in the eigenvectors
# Function to apply the Kolmogorov-Smirnov test and return non-Gaussian eigenvectors and their positions
check_gaussianity <- function(Dataset, alpha = 0.05) {
  non_gaussian <- list()  # To store non-Gaussian eigenvectors
  indices <- vector()     # To store corresponding indices
  
  corr_matrix <- scaled(Dataset = Dataset)
  
  eigenvectors <- eigen(corr_matrix, symmetric = T)$vectors
  
  for (i in 1:ncol(eigenvectors)) {
    eigenvector <- eigenvectors[, i]
    
    # Perform the KS test comparing the eigenvector to a normal distribution
    ks_test <- ks.test(eigenvector, "pnorm", mean = mean(eigenvector), sd = sd(eigenvector))
    
    # If the p-value is less than alpha, the eigenvector is considered non-Gaussian
    if (ks_test$p.value < alpha) {
      non_gaussian[[length(non_gaussian) + 1]] <- eigenvector
      indices <- c(indices, i)  # Save index
    }
  }
  return(list(non_gaussian_vectors = non_gaussian, positions = indices))
}

#Plot the inverse partecipation Ratio
IPR_fun <- function(Dataset){
  
  # Number of variables (N) and observations (M) for MP-Law parameters
  N <- ncol(Dataset)  # Number of variables
  M <- nrow(Dataset)  # Number of observations
  Q <- N / M          # Shape parameter (N / M)
  
  # Calculate the range for the Marchenko-Pastur distribution
  lambda_min <- (1 - sqrt(Q))^2
  lambda_max <- (1 + sqrt(Q))^2
  
  corr_matrix <- scaled(Dataset = Dataset)
  
  vectors <- eigen(corr_matrix)$vectors
  values <- eigen(corr_matrix, symmetric = T, only.values = T)$values
  
  IPRs <- vector(mode = "numeric", length = 98)
  
  for(i in 1:nrow(corr_matrix)){
  
    IPRs[i] <- sum(vectors[,i]^4)
  }
  
  non_gaussian <- check_gaussianity(Dataset = Dataset)
  
  # Highlight and plot non-Gaussian eigenvalues
  non_gaussian_positions <- non_gaussian[["positions"]]
  
  data_plot <- data.frame(eigenvalues = values, IPR = IPRs, gaussian_positions = rep("blue", times = 98))
  
  data_plot$gaussian_positions[non_gaussian_positions] <- "green" 

  plot(x = log10(data_plot$eigenvalues), y = log10(data_plot$IPR), type = "b",
       xlab = "log(Eigenvalues)", ylab = "log(Inverse Partecipation Ratio)", col = data_plot$gaussian_positions,
       main = paste("Plot of IPR against eigenvalues of ", deparse(substitute(Dataset))), xlim = c(-2,1.5), ylim = c(-2,-0.4))
  rect(log10(lambda_min), -2, log10(lambda_max), 0,col = rgb(0.5,0.5,0.5,1/4))
  abline(v = log10(lambda_min), col = "red")
  abline(v = log10(lambda_max), col = "red")
  
  legend("topright", legend = c("Associated to Non-Gaussian eigenvectors", "Associated to Gaussian eigenvectors"),
         col = c("green", "blue"), pch = c(16, 16), pt.cex = 2)
}
```

```{r}
IPR_fun(Dataset = Data_Pre)
```

```{r}
IPR_fun(Dataset = Data_Post)
```

The function mp_density computes the MP-Law and is used by subsequent functions.
The function generete_GOE generetes a Gaussian Orthogonal Ensamble and is used by subsequent functions.

This function produces the histogram of the eigenvalues with the MP-law and GOE histogram of eigenvalues overplotted 
```{r}
# MP distribution PDF
mp_density <- function(lambda, Q) {
  # Calculate the range for the Marchenko-Pastur distribution
  lambda_min <- (1 - sqrt(Q))^2
  lambda_max <- (1 + sqrt(Q))^2
  sqrt_part <- sqrt((lambda_max - lambda) * (lambda - lambda_min))
  return((1 / (2 * pi * Q * lambda)) * (sqrt_part))
}
generate_GOE <- function(m, n) {
  
  # Create a matrix with normally distributed random numbers
  GOE_matrix <- matrix(rnorm(n * m), nrow = m, ncol = n)
  
  return(GOE_matrix)
}

mp_plotting <- function(Dataset){
  
  # Scale and compute the correlation matrix
  corr_matrix <- scaled(Dataset = Dataset)
  
  # Get eigenvalues from the correlation matrix
  eigenval <- eigen(corr_matrix, symmetric = TRUE)$values
  
  # Plot histogram of eigenvalues from the dataset
  hist(eigenval, breaks = 200, probability = TRUE, 
       main = paste("Histogram of Eigenvalues with MP Law and GOE Matrix of", deparse(substitute(Dataset))),
       xlab = "Eigenvalues", col = "lightblue", ylim = c(0, 1))
  
  # Number of variables (N) and observations (M) for MP-Law parameters
  N <- ncol(Dataset)  # Number of variables
  M <- nrow(Dataset)  # Number of observations
  Q <- N / M          # Shape parameter (N / M)
  
  # Calculate the range for the Marchenko-Pastur distribution
  lambda_min <- (1 - sqrt(Q))^2
  lambda_max <- (1 + sqrt(Q))^2
  
  # Sequence for lambda and MP distribution values
  lambda_seq <- seq(lambda_min, lambda_max, length.out = 1000)
  mp_values <- sapply(lambda_seq, mp_density, Q = Q)
  
  # Add MP Law curve to the plot
  lines(lambda_seq, mp_values, col = "red", lwd = 2, lty = 2)
  
  # Generate a GOE matrix of size N x N (same size as correlation matrix)
  GOE_matrix <- generate_GOE(m = M, n = N) 
  
  GOE_corr <- scaled(GOE_matrix)
  
  GOE_eigen <- eigen(GOE_corr)$values
  
  # Flatten the GOE matrix to a vector to get all entries
  GOE_entries <- as.vector(GOE_eigen)
  
  # Add density curve for the GOE matrix
  lines(density(GOE_entries), col = "green", lwd = 2)
  
  # Add legend
  legend("topright", legend = c("Eigenvalues", "MP-Law", "GOE Entries"),
         col = c("lightblue", "red", "green"), lwd = 2, lty = c(1, 2, 1), fill = c(NA, NA, rgb(0, 1, 0, 0.5)))
}
```

This function shuffles the entries of the Dataset to build a null model.
```{r}
randomize_null <- function(Dataset) {

  n_rows <- nrow(Dataset)
  n_cols <- ncol(Dataset)
  
  # Flatten the Dataset into a vector
  Dataset_vector <- as.vector(Dataset)
  
  # Randomize the positions
  order <- sample(1:length(Dataset_vector), size = length(Dataset_vector), replace = FALSE)
  
  # Create a new randomized matrix with the same dimensions
  Dataset_randomized <- matrix(Dataset_vector[order], nrow = n_rows, ncol = n_cols)
  
  return(Dataset_randomized)
}
```

This function provides the heatmap of the non-gaussian eigenvectors over time
```{r}
###########
#Heatmaps with cropping
temporal_corr_cropped <- function(Dataset, window_size, step_width) {
  
  # Number of variables (N) and observations (M) for MP-Law parameters
  N <- ncol(Dataset)  # Number of variables
  M <- nrow(Dataset)  # Number of observations
  Q <- N / M          # Shape parameter (N / M)
  
  # Maximum and minimum eigenvalues based on MP-Law
  lambda_max <- (1 + sqrt(Q))^2
  lambda_min <- (1 - sqrt(Q))^2
  
  # Eigenvalue decomposition of the scaled dataset
  eigenvalues <- eigen(scaled(Dataset))$values
  
  # Find positions of eigenvalues outside the [lambda_min, lambda_max] range
  positions <- which(eigenvalues > lambda_max | eigenvalues < lambda_min)

  temporal_average <- matrix(0, nrow = length(positions), ncol = length(positions))
  
  rownames(temporal_average) <- positions
  colnames(temporal_average) <- positions
  
  cycles <- floor((nrow(Dataset) - (window_size+1))/step_width)
  
  for (i in 1:cycles) {
    # Adjust the sliding window calculation
    start_A <- i
    end_A <- start_A + window_size
    start_B <- start_A + step_width
    end_B <- start_B + window_size
    
    if(end_B > nrow(Dataset)) break
    
    Data_A <- Dataset[start_A:end_A,]
    Data_B <- Dataset[start_B:end_B,]
    
    # Center and compute correlation matrices
    Data_A <- scaled(Data_A)
    Data_B <- scaled(Data_B)
    
    # Eigen decomposition
    eigen_A <- eigen(Data_A, symmetric = T)
    eigen_B <- eigen(Data_B, symmetric = T)
    
    vectors_A <- eigen_A$vectors[, positions]
    vectors_B <- eigen_B$vectors[, positions]
    
    # Vectorized thresholding for eigenvectors A and B
    thresholds_A <- 1.8 * apply(vectors_A, 2, sd) + apply(vectors_A, 2, mean)
    vectors_A <- apply(vectors_A, 2, function(col, th) ifelse(abs(col) < th, 0, col), th = thresholds_A)
 
    thresholds_B <- 1.8 * apply(vectors_B, 2, sd) + apply(vectors_B, 2, mean)
    vectors_B <- apply(vectors_B, 2, function(col, th) ifelse(abs(col) < th, 0, col), th = thresholds_B)
    
    # Calculate temporal correlation matrix
    temporal <- t(vectors_A) %*% vectors_B
    temporal_average <- temporal_average + temporal
  }
  
  # Average the temporal correlations across all cycles
  temporal_average <- temporal_average / i
  
  # Final heatmap
  heatmap(temporal_average, Rowv = NA, Colv = NA, labRow = positions, labCol = positions,
          main = paste("Heatmap eigenvectors stability of", deparse(substitute(Dataset))),
          sub = paste("window size", window_size , "; skip = ", step_width))
}
```

